---
title: "Housing prices"
output: html_notebook
---

> Ask a home buyer to describe their dream house, and they probably won't begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition's dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence.
With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges you to predict the final price of each home.

[Source](https://www.kaggle.com/c/house-prices-advanced-regression-techniques)


*Goal*:<br>
Predict the sales price for each house. For each Id in the test set, you must predict the value of the SalePrice variable. 

*Metric*: <br>RMSE between log(predValue) and log(observered)´´


```{r,include=FALSE, echo=FALSE}
library(tidyverse)
library(mboost)
library(mgcv)
library(caret)
```

Myplan for this notebook is to get some reasonable performance (low overall error) applying a gadient descent boosting procedure with variable stability selection. The **mboost** function serves as workhorse and takes care of most of the variable selection for now. 
<br>
Okay. Lets first load the data and have a look at it:

```{r include=FALSE}
df.train <- read_csv("train.csv") %>% mutate(data = "train") # The training dataset
df.test <- read_csv("test.csv") %>% mutate(data = "test",SalePrice = NA) # The testing dataset
df <- bind_rows(df.train,df.test)

```

Next, lets format the variables considered for the modeling process. I make a couple of assumptions about which variables i include and are relevant.

```{r Precprocessing}
# Preformatting of variables
  # I think I will try those as continous linear features
  # * Longer street distance -> higher sale
  # * Greater LotArea -> higher sale
  
  df$Street[which(is.na(df$Street))] <- "No"
  df$Alley[which(is.na(df$Alley))] <- "No"
  
  # Missing values in the garage will be left missing and imputed later
  # df$GarageYrBlt
  # Same for lot frontage
  #df$LotFrontage
  
  # The type of utilities would influence my decision of buying anything
  # Neighboorhood could possibly be linked to a spatial structure assuming that closer neighb. have similar prices
  
  #df$Exterior1st[which(is.na(df$Exterior1st))] <- "No"
  #df$Exterior2nd[which(is.na(df$Exterior2nd))] <- "No"
  
  # No pool?
  df$PoolQC[which(is.na(df$PoolQC))] <- "No"
  # No Fence?
  df$Fence[which(is.na(df$Fence))] <- "No"
  # Misc Features
  df$MiscFeature[which(is.na(df$MiscFeature))] <- "No"
  # Fireplace
  df$FireplaceQu[which(is.na(df$FireplaceQu))] <- "No"
  # GarageFinish
  df$GarageType[which(is.na(df$GarageType))] <- "No"
  df$GarageFinish[which(is.na(df$GarageFinish))] <- "No"
  # Garage conditions
  df$GarageQual[which(is.na(df$GarageQual))] <- "No"
  df$GarageCond[which(is.na(df$GarageCond))] <- "No"
  
  df$MasVnrType[which(is.na(df$MasVnrType))]  <- "No"
  df$MasVnrType[which(is.na(df$MasVnrType))]  <- "No"
  df$MasVnrArea[which(is.na(df$MasVnrArea))]  <-  0 # Assuming 0 is reflectant of no areal
  # Basement
  df$BsmtCond[which(is.na(df$BsmtCond))] <- "No"
  df$BsmtExposure[which(is.na(df$BsmtExposure))] <- "None"
  df$BsmtFinType1[which(is.na(df$BsmtFinType1))] <- "No"
  df$BsmtFinType2[which(is.na(df$BsmtFinType2))] <- "No"
  # Electrical
  df$Electrical[which(is.na(df$Electrical))] <- "No"
  df$BsmtFinType2[which(is.na(df$BsmtFinType2))] <- "No"
  

df <- df %>% 
  mutate(
    MSSubClass = factor(MSSubClass), # Identifies the type of dwelling involved in the sale
    MSZoning = factor(MSSubClass), # zoning classification of the sale
    LotShape = factor(LotShape,levels = c("Reg","IR1","IR2","IR3"),ordered = T), # Ordered factor of irregularity
    LandContour = factor(LandContour, levels = c("Lvl","Bnk","HLS","Low"),ordered = T),# Ordered factor of the property flatness
    LandSlope = factor(LandSlope, levels = c("Gtl","Mod","Sev"),ordered = T), # "Landslope of property"
    Utilities = factor(Utilities), # Any utilities
    Neighborhood = factor(Neighborhood), # Good random intercept for now
    BldgType = factor(BldgType), # What kind of building
    HouseStyle = factor(HouseStyle), # What housestyle. Potentially nested in building type
    Remodelled = factor(ifelse(YearBuilt==YearRemodAdd,0,1)), # Was the house remodelled ?
    RoofStyle = factor(RoofStyle), RoofMatl = factor(RoofMatl), # How is the roof?
    # Create a factor whether there 0,1 or 2 conditions
    Conditions = factor( ifelse(Condition1=="Norm",0,ifelse(Condition2!="Norm",2,1)) ), # 
    Condition1 = factor( Condition1 ),
    Condition2 = factor( Condition2 ),
    # Exterior
    Exterior1st = factor( Exterior1st ),
    Exterior2nd =  factor( Exterior2nd ),
    # Other things
    CentralAir = factor(CentralAir),
    HeatingQC = factor(HeatingQC,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    Functional = factor(Functional,levels = rev(c("Typ","Min1","Min2","Mod","Maj1","Maj2","Sev","Sal")),ordered = T),
    PavedDrive = factor(PavedDrive),
    KitchenQual = factor(KitchenQual,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    FireplaceQu = factor(FireplaceQu,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    PoolQC = factor(PoolQC,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T), # Pool good
    GarageQual = factor(GarageQual,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    GarageCond = factor(GarageQual,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    GarageFinish = factor(GarageFinish,levels = c("No","Unf","RFn","Fin")),
    ExterQual = factor(ExterQual,levels = c("Po","Fa","TA","Gd","Ex"),ordered = T),
    ExterCond = factor(ExterCond,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    BsmtQual = factor(BsmtQual,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    BsmtCond = factor(BsmtCond,levels = c("No","Po","Fa","TA","Gd","Ex"),ordered = T),
    BsmtExposure = factor(BsmtExposure, levels = c("None","No","Po","Fa","TA","Gd","Ex"),ordered = T),
    BsmtFinType1 = factor(BsmtFinType1, levels = c("No","Unf","LwQ","Rec","BLQ","ALQ","GLQ" )),
    BsmtFinType2 = factor(BsmtFinType2, levels = c("No","Unf","LwQ","Rec","BLQ","ALQ","GLQ" )),
    
    Fence = relevel(factor(Fence),ref ="No") # Fence?
  )

# Check missing values
( x <- names( which( apply(df, 2, function(x) any(is.na(x))) ) ) )
# Here we missed some
(y <- names( which( apply(df, 2, function(x) any(!is.numeric(x)) ) )  ) ) # All non-numeric columns

# get those missing
df[,y[which(y %in% x)] ] <- df[,y[which(y %in% x)] ] %>% mutate_if(is.factor,as.character) %>%  replace(is.na(.),values =  "No")
df[,y[which(y %in% x)] ] <- df[,y[which(y %in% x)] ] %>% mutate_if(is.character,as.factor) # Convert back

# Did not work for those. Correct manually
df$GarageYrBlt <- as.numeric(as.character(df$GarageYrBlt))
df$LotFrontage <- as.numeric(as.character(df$LotFrontage))
df$BsmtFinSF1 <- as.numeric(as.character(df$BsmtFinSF1))
df$BsmtFinSF2 <- as.numeric(as.character(df$BsmtFinSF2))
df$BsmtUnfSF <- as.numeric(as.character(df$BsmtUnfSF))
df$GarageCars <- as.numeric(as.character(df$GarageCars))
df$GarageArea <- as.numeric(as.character(df$GarageArea))
df$TotalBsmtSF <- as.numeric(as.character(df$TotalBsmtSF))

df$SaleCondition <- factor(df$SaleCondition)
df$SaleCondition <- factor(df$SaleCondition)

# Rename first and second floor
df <- df %>% rename("firstFloor" = `1stFlrSF`,"secondFloor" = `2ndFlrSF`,"thirdPorch" = `3SsnPorch` )
```

# Model!
From here on lets build the model

```{r Model}
require(mboost)

# General
response = "SalePrice"
fam = Gaussian()

# Groups of variables 
var.linear <- c("Utilities","LotConfig","LandSlope","OverallQual","OverallCond","Conditions","Remodelled","MasVnrType","Foundation",
                "CentralAir","BsmtFullBath","BsmtHalfBath","FullBath","HalfBath","Bedroom","Kitchen","TotRmsAbvGrd","Fireplaces",
                "GarageType","GarageFinish","PavedDrive","BsmtFinType1","BsmtFinType2","Fence")
var.mono.dec <- c("LandContour","LotShape","Functional")
var.mono.inc <- c("ExterCond","ExterQual","BsmtQual","BsmtCond","BsmtExposure","HeatingQC","KitchenQual",
                  "FireplaceQu","GarageQual","GarageCond","PoolQC")
var.smooth <- c("LotFrontage","LotArea","YearBuilt","MasVnrArea","BsmtFinSF1","BsmtFinSF2","BsmtUnfSF","TotalBsmtSF","firstFloor","secondFloor","LowQualFinSF","GrLivArea","GarageYrBlt","GarageCars","GarageArea","WoodDeckSF","OpenPorchSF","EnclosedPorch","thirdPorch","ScreenPorch",
                "PoolArea","MiscVal","YrSold")
var.tree <- c("Street","Alley","Condition1","Condition2","BldgType","HouseStyle","RoofStyle","RoofMatl","Exterior1st","Exterior2nd","Heating","Electrical","MiscFeature","SaleType")
var.random <- c("MSZoning","SaleCondition","Neighborhood","MSSubClass") # To accompany for spatial and type structure

# Log-transform the response
pred$df.train[,response] <- log(pred$df.train[,response])

# Mean center / scale all continious predictors
pred$df.train[,c(var.smooth)] <- apply( pred$df.train[,c(var.smooth)], 2, function(y) { scale(y,center = TRUE, scale = F) } )
pred$df.test[,c(var.smooth)] <- apply( pred$df.test[,c(var.smooth)], 2, function(y) { scale(y,center = TRUE, scale = F) } )

# Boosting control options
bctrl = boost_control(mstop = 5000,nu = 0.001,risk = "inbag", trace = TRUE)
# Set step-length factor  to about 0.001 for maximal robustness  
# All base-learners were made comparable by centering predictors at the beginning and using the same degrees of freedom for all base-learners 

## Define the base learners ##
base.linear <- paste("bols(", var.linear, ",intercept = FALSE, df = 1)", sep = "", collapse = " + ")
base.smooth <- paste("bbs(", var.smooth, ",knots =4, center = TRUE, df = 1)", sep = "", collapse = " + ")
base.mono.dec <- paste("bmono(", var.mono.dec, ",constraint = \"decreasing\", df = 1)", sep = "", collapse = " + ")
base.mono.inc <- paste("bmono(", var.mono.inc, ",constraint = \"decreasing\", df = 1)", sep = "", collapse = " + ")
base.tree <- paste("btree(", var.tree, ")", sep = "", collapse = " + ")
base.random <- paste("brandom(", c(var.random), ")", sep = "", collapse = " + ")

# Construct formulas
f = as.formula( paste(response, " ~ ",
                      #base.linear, "+",
                      base.smooth, "+",
                      base.mono.dec, "+",
                      base.mono.inc, "+",
                      base.tree, "+",
                      base.random, collapse = " + "  ) )

# Start model
mb.full <- mboost(formula = f, control = bctrl,data=pred$df.train, family=fam)

# Get a first prediction
val <- exp( mboost::predict.mboost(mb.full,newdata = pred$df.test,type = "response") )

write.csv(data.frame(Id = pred$df.test$Id, val),"predictions1.csv",header=TRUE)

### -- Crossvalidating with OOB all models -- ###
# Early stopping using Cross-validation to prevent potential overfitting the model
crossStop <- function(mb1,B=10){
  cv10f <- cv(model.weights(mb1), type = "kfold",B=B,prob = 0.5, strata = pred[mb1$rownames,"bcr"] ) 
  cvm <- cvrisk(mb1, folds = cv10f) # Parallel optimization
  return(cvm)
}
# Crossvalidate
cvm.local <- crossStop(mb.local)
mb.local[mstop(cvm.local)] # Set the max number of boosting iterations



```

